Peer Assignment 1 - Edward Pittman
========================================================



```{r, echo=TRUE}
setwd("/Users/edward/Documents/R/Rcourse2014May")   #  set working directory
act1 = read.csv("activity.csv", header = TRUE)    #  read activity.csv file
act2 <- na.omit(act1)          # act2 excludes missing values

#install.packages("plyr")   #installs plyr package
library(plyr)

#creates summary data frame act3 sums steps by day
act3 <- ddply(act2,~date,summarize,stepsum=sum(steps))   

#prints histogram of steps
hist(act3$stepsum, breaks=15,
     main="Histogram of Total Number of Steps Per Day",
     ylab="number of days",
     xlab="total number of steps per day")

meanstep <- mean(act3$stepsum)   # calculate mean steps per day
medstep <- median(act3$stepsum)    # calulate median steps per day
```
The mean total number of steps per day is `r meanstep`. The median number of total steps per day is `r medstep`.
```{r, echo=TRUE}
# creates summary data frame act4 grouped by time of day
act4 <- ddply(act2,~interval,summarize,stepmean=mean(steps))   

plot.ts(act4$stepmean, type="l")     #plots mean steps by time of day
tmp1 <- which.max(act4$stepmean)   #calculates time of day for max steps
maxtime <- act4[tmp1,1]
```
The 5-minute interval with the maximum number of steps is `r maxtime`.
```{r, echo=TRUE}
# creates act5 by merging act1 and act4 by interval to impute missing values
act5 <- merge(act1,act4,by="interval",all=TRUE)
act5$ct1 <- ifelse(!is.na(act5$steps),0,1)  #1 if na
ctna <- sum(act5$ct1) #count of na's
act5$steps2 <- ifelse(act5$ct1==1,act5$stepmean,act5$steps)

#creates summary data frame act5 sums steps by day
act6 <- ddply(act5,~date,summarize,stepsum=sum(steps2))   



#prints histogram of steps
hist(act6$stepsum, breaks=15,
     main="Histogram of Total Number of Steps Per Day 
     \n missing values imputed",
     ylab="number of days",
     xlab="total number of steps per day")
meanstepm <- mean(act6$stepsum)   
# calculate mean steps per day w imputed missing
medstepm <- median(act6$stepsum)  
# calulate median steps per day-imputes missing
```

The mean and median number of total steps per day with missing values imputed are `r meanstepm` and `r medstepm` respectively. The mean values are the same with and without imputed data for missing values. The median with imputed data is slightly higher than the median excluding missing values.

```{r, echo=TRUE}
# creates categorical variable daytype indicating weekday or weekend
act5$x1 <- as.Date(act5$date)
act5$x2 <- weekdays(act5$x1)
act5$daytype <- "weekday"
act5$daytype <- ifelse(act5$x2=="Saturday","weekend",act5$daytype)
act5$daytype <- ifelse(act5$x2=="Sunday","weekend",act5$daytype)

# creates summary data frame act5 grouped daytype, then by time of day
act7 <- ddply(act5,.(daytype, interval),summarize,stepmean=mean(steps2))
library(ggplot2)
actwkd <- subset(act7,daytype=="weekday")
actwke <- subset(act7,daytype=="weekend")

plot.ts(actwkd$stepmean, type="l", main="weekday")

plot.ts(actwke$stepmean, type="l", main="weekend") 

```







